% This is a template for your written document.
%
% To compile using latexmk on the command line, run the following: 
% latexmk -pdf main.tex

\documentclass[12pt]{article}
\usepackage{setspace}
\singlespace
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{graphicx}

\title{\textbf{Reflective AI Feedback for Personal Productivity}}
\author{Saidamir Osimov}

\begin{document}

\maketitle

\section{Introduction}
Productivity systems are everywhere, they have become deeply embedded in the daily life of millions of people, whether they work
 in the office, schools, or a project. They help maintain consistency and track tasks and goals. By reviewing someone's schedule 
 and task list in their productivity app, you can learn a great deal about them, such as how busy they are throughout the week, how 
 productive they are each day, and how many meetings and tasks they have. Task managers, digital planners, goal tracking dashboards,
  habit-building apps, users now have multiple tools to organize their life and structure their time. All these platforms are great
   at dividing the work into actionable small steps, setting deadlines and monitoring completion rates across different horizons.
     While they are great at tracking and supporting execution, they rarely foster insight. Users can rigorously record what they do, 
     but they receive little to no support in understanding how they work.

\subsection{The Problem of Modern Productivity Systems}
At the rate of tasks and pressure of commitments, individuals face cognitive overload. When, tasks are completed, postponed, or 
rescheduled, it is often done so without any coherent decisions making behind them. Productivity tools generate data about the
 person more than ever before, but they provide minimal awareness about behavioral patterns, motivational cycles, or attention
  shifts. Tracking an activity is not the same as understanding it. Without systems for reflection, productivity tools are more like 
  sophisticated record keepers rather than instruments for self-improvement.

\subsection{ The Need for Reflective Feedback }
Reflection plays an important role in cognitive growth and sustained performance. Research in the field of cognitive science 
consistently shows that individuals improve not just by completing tasks, but by investigating how and why they approach them. 
In particular, a 2025 study, “Effect of feedback-integrated reflection on deep learning of medical students” \cite{Maqsood2025}, 
found that combining structured reflection with feedback, caused significantly higher learning gains (approximately 69\% compared 
to 29\%). The study also highlights how reflective feedback reinforces metacognitive activity, enabling learners to monitor thinking
 and adjust behaviors effectively. When individuals look back and analyze patterns in their performance and attention, they develop 
 awareness that supports more goal-aligned action.

Although, reflective practice is widely supported by research, most productivity platforms remain focused on task completion and 
execution rather than metacognitive development(awareness and understanding your own thought process). Although users can record 
activities and track results, they are rarely encouraged to analyze their patterns or get personalized feedback. The absence of
 smart reflection systems shows a major disconnect between what cognitive science knows and what productivity tools actually offer.

Recent progress in artificial intelligence, particularly large language models allow productivity systems to go beyond task 
management, toward cognitive support. Unlike traditional software, artificial intelligence can analyze large volumes of data and
 detect patterns across time and interpret contextual details in user generated logs. This allows the system to make sense of 
 information rather than just collecting it. Large language models are capable of identifying recurring themes in work behavior, 
 demonstrating inconsistencies between stated goals and actual actions, and generating reflection feedback in natural language.
  Artificial intelligence can help users to interpret their own patterns of focus, procrastination, prioritization, and energy
   allocation. In this sense artificial intelligence functions as a mechanism for cognitive augmentation. By letting intelligent 
   systems handle pattern recognition and reflection, users can develop self-awareness without having to do all the mental heavy
    lifting themselves.


\subsection{	Introducing Newton}
Newton is a productivity system designed to have artificial intelligence driven reflective feedback directly into the everyday 
workflow management. Built on top of the Supermind Design concept introduced by MIT\cite{10.1145/3643562.3672611}, the platform
 encourages creative thinking, and intentional execution rather than passive task tracking. The core functionality of Newton lies
 in its analysis of user activity through structured change logs. 
As users edit tasks, mark them complete and update goals, these changes are then captured and interpreted by an integrated 
artificial intelligence component. The systems analyzes the log to detect recurring patterns, and inconsistencies. Based on
 this analysis, Newton generates personalized reflective insights that help users to zoom out of the daily routine and evaluate
  how their actions are aligned with their long term goals. Rather than simply reporting what has been completed, the system 
  highlights how your work develops over time and where you might need to adjust.
This project explores how integrating AI-generated reflective feedback into productivity systems can enhance user awareness, 
improve workflow efficiency, and support more intentional decision-making.

\section{Background}
The Supermind Design method is humans and computers working together to generate better solutions to the problems. Rather than 
replacing human thinking, these systems combine human creativity and judgement with computational power of the machines. Also, the
 Supermind Design is an approach to problem solving that combines big picture thinking with creative and intentional actions. Rather 
 than focusing on isolated tasks, this method encourages individuals to view components to treat problems as parts of a larger, 
 connected whole rather than standalone issues to fix.  A key part of this approach is looking at things from different angles, and 
 considering both the small details and the bigger picture before making any decisions. By stepping back before diving into the 
 details users can better spot root problems, and hidden issues that might be missed with a narrow focus. Supermind Design also
  promotes iterative reflection rather than strictly following a plan. Reframing plays a key role in this method, encouraging users 
  to question assumptions and find alternative solution paths. This framework blends human intuition with structured reasoning.

\subsection{Supermind Design in Newton}
Newton applies the Supermind Design principles by combining human decision making with artificial intelligence generated insights. 
Instead of operating only as a task manager the system is designed to support the collaboration between the user and the artificial
 intelligence systems, that gives reflective feedback and analyzes the behavioral patterns.
By analyzing the logs, the artificial intelligence component is able to help users see the patterns over time in their day-to-day 
work. This is the reflection of Supermind Design: combining humans with computational analysis to get better insights on productivity
 and workflow.

\subsection{AI as Collaborative Agent}
Artificial intelligence does not replace humans, it enhances them. The invention of the calculator did not replace the need for
 mathematicians, or math ability, it sped it up. An LLM does not replace thinking, it improves and supports reasoning. Newton is
  not doing the work for the user, it is helping them see the patterns in their own behavior and empowering them to look at their
   workflow from another perspective. That is augmentation. Artificial intelligence expands cognitive capacity. 

A tool waits for the instructions, Newton is more like a partner that: 1) Observes 2) Suggests 3) Reflects. Traditional productivity
 apps are passive, Newton on the other hand is reflective. That is the difference. If I ask Siri a question, then it will give me 
 an answer. I use Newton, it analyzes my behavior through the progress logs and tells me: “You tend to overcommit midweek” or “You
  abandon long-term tasks after 5 days”.





\section{Related Works}
Researchers have intensely studied how artificial intelligence systems can support task management and everyday work activities. 
Work by D. Trippas 
and colleagues \cite{10.1145/3295750.3298934} categorizes common workplace tasks to help design intelligent assistants that better 
understand the goals of the users and their activities. The study demonstrates that there are common patterns in work activities and
 tasks, most commonly they involve multiple step tasks, goals and contextual factors.
Other research \cite{KHAOKAEW2022102905} on future digital assistants in the workplace suggests that users want tools that are not
 just reactive tools but also adaptive supporters of complex work processes.
Traditional digital assistants respond to commands and perform tasks. However, these studies indicate that the workers increasingly
 expect the assistants to understand context and support planning and decision making. On the same topic, Newton analyzes user task
  history and change logs to generate reflective feedbacks. Instead of simply following the commands, it aims to help users 
  understand the patterns in their workflow and support them in more informed decision making.

Recent research frames artificial intelligence as a tool for augmented human cognition, rather than replacing human decision making.
 What that means, is that nowadays artificial intelligence is used as a help humans think better and help them in decision making.
  For example, Supermind Ideator developed by Thomas W. Malone and collaborators \cite{10.1145/3643562.3672611} showcases how 
structured human and artificial intelligence collaboration can improve creative problem solving. By offering support such as 
rephrasing problems, prompting new ideas, and guiding collaboration, the system helped users come up with significantly more
 creative ideas than when working alone or using regular AI tools. Similarly, research on the ExtendAI framework by Leon Reicherts 
 and colleagues \cite{10.1145/3706598.3713295} shows that 
AI systems are more effective when they extend and support users' reasoning processes rather than simply providing recommendations.
 Their findings suggest that AI tools that integrate with users' own decision-making strategies lead to better engagement and 
 slightly improved outcomes. Research by Xiaotong Xu \cite{10.1145/3706598.3713649} also shows that adding AI to creative 
 workflows can help users generate more ideas and explore problems more deeply.
Together, these studies point to a bigger shift in how artificial intelligence systems are being designed. The goal is always
 to support human thinking, not replace it. Newton takes this same approach by integrating AI into everyday productivity workflows,
  using data and task history to help users reflect on and improve their work habits.




%\begin{figure}[b] puts the image to the (closest) bottom of page
%\begin{figure}[t] puts the image to the (closest) top of page
%\begin{figure}[h] allows the image to float, putting it roughly wherever it was placed in the text.

\section{Methodology}
Newton is developed as a reflective productivity system that tracks not only what users accomplish, but also how they accomplish 
it over time. Instead of being a typical task manager that prioritizes execution and completion, Newton is more focused on pattern 
recognition and behavioral analysis. This is done to shift the usual task tracking toward deeper self-awareness. The system analyzes
 user activity, aiming to help individuals understand their habits, inconsistencies between planned goals and actual behavior, and
  decision-making. The system is not intended to replace human thinking or decision-making, it is rather made to support it. 
  Artificial intelligence in Newton is used as a form of cognitive augmentation, helping users understand the information about 
  their own workflow that would otherwise be difficult to track manually over time. By showcasing behavioral patterns and providing 
  reflective feedback to users, Newton encourages them to step back and evaluate their work and make more intentional decisions on 
  how to spend energy and time on tasks more efficiently.

A key feature of Newton's approach lies in its ability to record any change made in the platform regarding tasks and user activity 
over time. The system continuously logs any changes made in the platform in the Progress panel. These changes include events such as 
task creation, task completion, deletion, rescheduling, and edits along with their timestamps. Instead of treating each task as a 
single isolated action, these logs create a continuous record of how a user behaves and interacts with their own workflow, reflecting
 the real behavior. For instance, if a user frequently reschedules a task, it may be an indicator of overcommitment. Similarly, 
 repeated edits may suggest uncertainty or shifting priorities. By collecting this information about the user, Newton is able to go 
 beyond simple tracking and begin identifying behavioral patterns. This provides a foundation for generating reflective insights for
  users to better understand how they perform over time.
\section{Conclusion}

\bibliographystyle{acm}
\bibliography{bibliography.bib}

\end{document}
